---
title: "practice"
author: "Kimon"
date: "2/7/2020"
output: 
	html_document:
		toc: yes
		toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)

DT <- as.data.table(mtcars)
DT[, am := as.character(am)]
DT[am == '1', am := 'manual']
DT[am == '0', am := 'automatic']
DT[, vs := as.character(vs)]
DT[vs == '1', vs := 'V']
DT[vs == '0', vs := 'straight']
DT[, am := as.factor(am)]
DT[, vs := as.factor(vs)]

head(DT)
```

# Descriptive

```{r}
summary(DT)
```

## Centre

```{r}
# Ignore outter-most values,
print(c( 'median' = median(DT$mpg),
				 'mean' = mean(DT$mpg), 
				 '5% trimmed' = mean(DT$mpg, trim=0.025),
				 'weighted' = weighted.mean(DT$mpg, rnorm(nrow(DT))), # random weights, dataset doesn't have any
				 'geometric' = exp(mean(log(DT$mpg))),
				 'harmonic' = 1/(mean(1/DT$mpg))
				 ))
```

## Spread

```{r}
print(c( 'min' = min(DT$mpg),
				 'max' = max(DT$mpg),
				 'range' = range(DT$mpg),
				 'quantiles' = quantile(DT$mpg, c(0.25, 0.5, 0.75)),
				 'variance' = var(DT$mpg),
				 'stdev' = sd(DT$mpg),
				 'sterr' = sd(DT$mpg) / sqrt(length(DT$mpg)),
				 'coeffvar' = sd(DT$mpg) / mean(DT$mpg), # ... * 100
				 '95% ci' = c( mean(DT$mpg) - 1.96 * sd(DT$mpg) / sqrt(length(DT$mpg)),  # mean +- Tstat * sterr
				 							 mean(DT$mpg) + 1.96 * sd(DT$mpg) / sqrt(length(DT$mpg)) ),  # 99% 2.58,  normal-ish only
				 '95% boot ci' = quantile(vapply(1:1000, 
				 																function(i){ 
				 																	mean(sample(DT$mpg, nrow(DT), replace=TRUE))
				 																}, numeric(1)),
				 												 c(0.025, 0.975))
))
```

## Viz

### One var

```{r}
library(ggplot2)

ggplot(DT, aes(x=1:nrow(DT), y=mpg)) + 
	geom_point() + 
	geom_line() + 
	geom_smooth(method='lm')
```

```{r}
ggplot(DT, aes(x="", y=mpg)) + 
	geom_boxplot()
```

```{r}
ggplot(DT, aes(x="", y=mpg)) + 
	geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))
```

```{r}
ggplot(DT, aes(x=mpg)) + 
	geom_histogram()
```

```{r}
ggplot(DT, aes(x=am, fill=am)) + 
	geom_bar()
```

```{r}
ggplot(DT, aes(x=am, fill=am)) + 
	geom_bar() +
	coord_polar(theta='y')
```

```{r}
ggplot(DT, aes(x=factor(""), fill=am)) + 
	geom_bar() +
	coord_polar(theta='y')
```

### Two vars

```{r}
ggplot(DT, aes(x=hp, y=mpg)) + 
	geom_point() + 
	geom_line() + 
	geom_smooth(method='lm')
```

```{r}
ggplot(DT, aes(x=hp, y=mpg)) + 
	geom_point() + 
	geom_density_2d()
```

### More vars

```{r}
ggplot(DT, aes(x=hp, y=mpg)) + 
	facet_wrap(~ cyl, scales="free") +
	geom_point() + 
	geom_smooth(method='lm')
```

```{r}
ggplot(DT, aes(x=hp, y=mpg, size=cyl)) + 
	geom_point(alpha=0.4)
```

### All the vars, pairwise

```{r, fig.height=10, fig.width=10}
library(GGally)

ggpairs(DT)
```

# Check for normality

```{r}
ggplot(DT, aes(x=qsec)) + 
	geom_density()
```

```{r}
qqline(DT$qsec, lty=2)
```

### Shapiro-Wilkinson test

N < 10 won't detect almost anything
N > 30 will detect almost everything

```{r}
print( shapiro.test(DT$qsec) )
```

### Kolmogorov-Smirnov (K-S)

Continuous distribution, does not like repeated values.

Compares against a specific distribution. Do NOT use for testing normality.

```{r}
print( ks.test(DT$hp, pnorm) )    # pnorm() : normal distribution
```

### More in package `nortest`

`nortest` package:

- Lilliefors test, based on K-S
- Anderson-Darling test, more accurate than lillie.


# Hypothesis testing - Continuous variables

## T-test - Mean of a normal distribution

```{r}
t.test(DT$qsec, mu=0)	# Test if mean is sig different from 0.
```

## Difference between means of two normal distributions - T-test

### Different variances in unpaired data (Student's)

"Welch 2-sample t-test", DOF ~< n-2

```{r}
# Different variances
t.test(DT$drat, DT$wt)	# Test if difference of means is sig different from 0.
```

### Equal variances in unpaired data

(Check variance equality using F-test or Levene's test)

"2-sample t-test", DOF == n-2

```{r}
t.test(DT$wt, DT$drat, var.equal = TRUE)	# Test if means sig different from each other.
```

### Matched value pairs

```{r}
t.test(DT$wt, DT$drat, paired=TRUE)	# # Test if means sig different from each other.
```

Or as a 1-sample test:

```{r}
t.test(DT$wt - DT$drat, mu=0)	 # Test if difference of means is sig different from 0.
```

## Difference between means of two non-normal distributions

Transform data to normal?

### Mann-Whitney / unpaired Wilcoxon rank sum








