---
title: "COVID19"
author: "Kimon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

# Libraries

```{r message=FALSE}
library(httr)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(data.table)
library(maps)

options(scipen=2)
```

Source: https://opendata.ecdc.europa.eu/covid19/

# Load current data

```{r}
# Load data
###########

## Download latest data
# GET("https://opendata.ecdc.europa.eu/covid19/agecasesnational/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/agecasesnational.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/agecasesubnational/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/agecasesubnational.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/casedistribution.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/hospitalicuadmissionrates/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/hospitalicuadmissionrates.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/movementindicators/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/movementindicators.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/nationalcasedeath/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/nationalcasedeath.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/subnationalcasedaily/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/subnationalcasedaily.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/subnationalcaseweekly/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/subnationalcaseweekly.csv', overwrite = TRUE))
# GET("https://opendata.ecdc.europa.eu/covid19/testing/csv", authenticate(":", ":", type="ntlm"), write_disk('D:/Documents/covid/testing.csv', overwrite = TRUE))

# Import data.
# Don't need the cruiseship.
agecasenat <- fread('D:/Documents/covid/agecasesnational.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
agecasesubnat <- fread('D:/Documents/covid/agecasesubnational.csv') # [country != 'Cases_on_an_international_conveyance_Japan',] # Still blank
casedist <- fread('D:/Documents/covid/casedistribution.csv')[countriesAndTerritories != 'Cases_on_an_international_conveyance_Japan', ]
hosprate <- fread('D:/Documents/covid/hospitalicuadmissionrates.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
movement <- fread('D:/Documents/covid/movementindicators.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
natcasedeath <- fread('D:/Documents/covid/nationalcasedeath.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
subnatcaseday <- fread('D:/Documents/covid/subnationalcasedaily.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
subnatcaseweek <- fread('D:/Documents/covid/subnationalcaseweekly.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]
testing <- fread('D:/Documents/covid/testing.csv')[country != 'Cases_on_an_international_conveyance_Japan', ]

# Verify that the data format wasn't changed again
stopifnot(names(agecasenat) == c('country', 'country_code', 'year_week', 'age_group', 'new_cases', 'population', 'rate_14_day_per_100k', 'source'))
stopifnot(names(agecasesubnat) == c('Test'))
stopifnot(names(casedist) == c('dateRep', 'year_week', 'cases_weekly', 'deaths_weekly', 'countriesAndTerritories', 'geoId', 'countryterritoryCode', 'popData2019', 'continentExp', 'notification_rate_per_100000_population_14-days'))
stopifnot(names(hosprate) == c('country', 'indicator', 'date', 'year_week', 'value', 'source', 'url'))
stopifnot(names(movement) == c('col0', 'country', 'country_code', 'geo_id_final', 'region', 'subnational_cases_7', 'subnational_cases_14', 'subnational_population', 'subnational_rate_14', 'national_cases_7', 'national_population', 'national_testing_rate', 'national_positivity_rate', 'subnational_testing_data', 'positivity_rate_combined', 'testing_rate_combined', 'colour', 'week'))
stopifnot(names(natcasedeath) == c('country', 'country_code', 'continent', 'population', 'indicator', 'weekly_count', 'year_week', 'rate_14_day', 'cumulative_count', 'source'))
stopifnot(names(subnatcaseday) == c('country', 'region_name', 'nuts_code', 'date', 'rate_14_day_per_100k', 'source'))
stopifnot(names(subnatcaseweek) == c('country', 'region_name', 'nuts_code', 'year_week', 'rate_14_day_per_100k', 'source'))
stopifnot(names(testing) == c('country', 'country_code', 'year_week', 'level', 'region', 'region_name', 'new_cases', 'tests_done', 'population', 'testing_rate', 'positivity_rate', 'testing_data_source'))


# Countries to highlight
########################

homecountry = 'Austria'
# Selected countries
family_countries <- c('Austria', 'Italy', 'Greece', 'Luxembourg', 'Germany')
neighbour_countries <- c('Switzerland', 'Slovakia', 'Slovenia', 'Czechia', 'Hungary')
other_countries <- c('United_Kingdom', 'Spain', 'France', 'Belgium', 'United_States_of_America', 'Sweden', 'South_Korea')


# Make dates workable and uniform
#################################

# str(agecasenat)
##str(agecasesubnat)
# str(casedist)
casedist[, dateRep := dmy(dateRep)]
# str(hosprate)
hosprate[, date := as.Date(date)]
hosprate[, year_week := sub('W', '', year_week)]
# str(movement)
# str(natcasedeath)
# str(subnatcaseday)
subnatcaseday[, date := as.Date(date)]
# str(subnatcaseweek)
subnatcaseweek[, year_week := sub('W', '', year_week)]
# str(testing)
testing[, year_week := sub('W', '', year_week)]


# Map outlines
##############

seablue <- '#0077FF'
world_map <- as.data.table(map_data('world'))

## Code template
# ggplot(world_map, aes(x=long, y=lat, group=group)) +
#   geom_polygon(fill='white', colour='black', size=0.01) +
#   coord_cartesian(ylim=c(-60, 80), xlim=c(-180,180)) + # Crop giant Antarctica and empty Arctic
#   theme(panel.background=element_rect(fill=seablue),
#         panel.grid=element_line(colour='white', size=0.1))


# Sanitize country names

# Map data names allow spaces, ECDC uses underscores.
# Spaces are nice, but ECDC tables out-number the coordinates table, so it makes sense to keep the ECDC convention.
# Also, ECDC groups some of these territories together, whereas the map doesn't (except for the Virgin Islands, where it is the other way around)/
world_map[, region := gsub(' ', '_', region,)]

# More mismatched spellings.
world_map[region=='Antigua', region := 'Antigua_and_Barbuda']
world_map[region=='Bonaire', region := 'Bonaire, Saint Eustatius and Saba']
world_map[region=='Sint_Eustatius', region := 'Bonaire, Saint Eustatius and Saba']
world_map[region=='Saba', region := 'Bonaire, Saint Eustatius and Saba']
## world_map[region=='Virgin_Islands', region := 'British_Virgin_Islands']       # I could try to sum the ECDC data, but more trouble than it's worth. 
## world_map[region=='Virgin_Islands', region := 'United_States_Virgin_Islands'] # Fixing the report structure is the priority.
world_map[region=='Brunei', region := 'Brunei_Darussalam']
world_map[region=='Republic_of_Congo', region := 'Congo']
world_map[region=='Ivory_Coast', region := 'Cote_dIvoire']
world_map[region=='Curacao', region := 'Curaçao']
world_map[region=='Czech_Republic', region := 'Czechia']
world_map[region=='Falkland_Islands', region := 'Falkland_Islands_(Malvinas)']
world_map[region=='Guinea-Bissau', region := 'Guinea_Bissau']
world_map[region=='Vatican', region := 'Holy_See']
world_map[region=='Micronesia', region := 'Micronesia_(Federated_States_of)']
world_map[region=='Nevis', region := 'Saint_Kitts_and_Nevis']
world_map[region=='Saint_Kitts', region := 'Saint_Kitts_and_Nevis']
world_map[region=='Grenadines', region := 'Saint_Vincent_and_the_Grenadines']
world_map[region=='Saint_Vincent', region := 'Saint_Vincent_and_the_Grenadines']
world_map[region=='Macedonia', region := 'North_Macedonia']
world_map[region=='Timor-Leste', region := 'Timor_Leste']
world_map[region=='Trinidad', region := 'Trinidad_and_Tobago']
world_map[region=='Tobago', region := 'Trinidad_and_Tobago']
world_map[region=='Turks_and_Caicos_Islands', region := 'Turks_and_Caicos_islands']
world_map[region=='UK', region := 'United_Kingdom']
world_map[region=='Tanzania', region := 'United_Republic_of_Tanzania']
world_map[region=='USA', region := 'United_States_of_America']
world_map[region=='Swaziland', region := 'Eswatini']

# Countries that still don't match
print( data.frame(Only_in_the_data = setdiff(
  unique(casedist$countriesAndTerritories), 
  unique(world_map$region)) ) )
print( data.frame(Only_in_the_map = setdiff(
  unique(world_map$region), 
  unique(casedist$countriesAndTerritories)) ) )

 
# Add coordinates to the data tables
####################################

mapify <- function(dt, country1='country', country2='region') {
  merge(dt, world_map, by.x=country1, by.y=country2, all.x=TRUE, all.y=FALSE, allow.cartesian=TRUE)
}

# Code templates
## Don't merge data with the coordinates yet. Leave it until after any calculations.
# agecasenat <- mapify(agecasenat)
# casedist <- mapify(casedist, 'countriesAndTerritories')
# hosprate <- mapify(hosprate)
# natcasedeath <- mapify(natcasedeath)
# testingnat <- mapify(testing[level == 'national',])
## These tables include subregional detail. Neither the data nor the coordinates have subregional info for all countries. So, merging on subregions is problematic, and merging on countries is not sensible for the kind of data.
# testingsubnat <- testing[level != 'national',]
## movement
## subnatcaseday
## subnatcaseweek

# Colours
#########
case_col = '#FF0000'
death_col = '#0055FF'
hosp_col = '#555500'
icu_col = '#995500'
test_col = '#005500'

# Normalization
popnorm= 1e6
```



# Status

```{r}
endday <- unique(casedist[dateRep == max(dateRep), dateRep])
startday <- unique(casedist[dateRep == min(dateRep), dateRep])
endweek <- unique(casedist[year_week == max(year_week), year_week])
startweek <- unique(casedist[year_week == max(year_week), year_week])
```

Based on the latest available date in the table. Some countries may be slower at releasing/entering their data.

## World's current worst

### By cases

```{r}
# Worse 10 (or more) by weekly cases
subDT <- casedist[dateRep==endday & !is.na(cases_weekly), 
                  .(countriesAndTerritories, dateRep, cases_weekly, population_normalized = cases_weekly / popData2019 * popnorm)] [order(cases_weekly, decreasing=TRUE), ]

n <- max(10, nrow(subDT[sapply(subDT$cases_weekly, 
                               function(x) { 
                                 all.equal(subDT$cases_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))

print( head(subDT, n) )

# Worse 10 (or more) by weekly cases normalized
setorder(subDT, -population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```

### By deaths

```{r}
# Worse 10 (or more) by weekly deaths
subDT <- casedist[dateRep==endday & !is.na(deaths_weekly), 
                  .(countriesAndTerritories, dateRep, deaths_weekly, population_normalized = deaths_weekly / popData2019 * popnorm)] [order(deaths_weekly, decreasing=TRUE), ]

n <- max(10, nrow(subDT[sapply(subDT$deaths_weekly, 
                               function(x) { 
                                 all.equal(subDT$deaths_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )

# Worse 10 (or more) by weekly deaths normalized
setorder(subDT, -population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```

## Europe's current worst

### By cases

```{r}
# Worse 10 (or more) by weekly cases
subDT <- casedist[dateRep==endday & !is.na(cases_weekly) & continentExp == 'Europe', 
                  .(countriesAndTerritories, dateRep, cases_weekly, population_normalized = cases_weekly / popData2019 * popnorm)] [order(cases_weekly, decreasing=TRUE), ]

n <- max(10, nrow(subDT[sapply(subDT$cases_weekly, 
                               function(x) { 
                                 all.equal(subDT$cases_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))

print( head(subDT, n) )

# Worse 10 (or more) by weekly cases normalized
setorder(subDT, -population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```

### By deaths

```{r}
# Worse 10 (or more) by weekly deaths
subDT <- casedist[dateRep==endday & !is.na(deaths_weekly) & continentExp == 'Europe', 
                  .(countriesAndTerritories, dateRep, deaths_weekly, population_normalized = deaths_weekly / popData2019 * popnorm)] [order(deaths_weekly, decreasing=TRUE), ]

n <- max(10, nrow(subDT[sapply(subDT$deaths_weekly, 
                               function(x) { 
                                 all.equal(subDT$deaths_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )

# Worse 10 (or more) by weekly deaths normalized
setorder(subDT, -population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```

## Europe's current best

### By cases

```{r}
# Worse 10 (or more) by weekly cases
subDT <- casedist[dateRep==endday & !is.na(cases_weekly) & continentExp == 'Europe', 
                  .(countriesAndTerritories, dateRep, cases_weekly, population_normalized = cases_weekly / popData2019 * popnorm)] [order(cases_weekly, decreasing=FALSE), ]

n <- max(10, nrow(subDT[sapply(subDT$cases_weekly, 
                               function(x) { 
                                 all.equal(subDT$cases_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))

print( head(subDT, n) )

# Worse 10 (or more) by weekly cases normalized
setorder(subDT, population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```

### By deaths

```{r}
# Worse 10 (or more) by weekly deaths
subDT <- casedist[dateRep==endday & !is.na(deaths_weekly) & continentExp == 'Europe', 
                  .(countriesAndTerritories, dateRep, deaths_weekly, population_normalized = deaths_weekly / popData2019 * popnorm)] [order(deaths_weekly, decreasing=FALSE), ]

n <- max(10, nrow(subDT[sapply(subDT$deaths_weekly, 
                               function(x) { 
                                 all.equal(subDT$deaths_weekly[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )

# Worse 10 (or more) by weekly deaths normalized
setorder(subDT, population_normalized, na.last=TRUE)

n <- max(10, nrow(subDT[sapply(subDT$population_normalized, 
                               function(x) { 
                                 all.equal(subDT$population_normalized[min(10, nrow(subDT))], x)
                               }) == TRUE, ]))
print( head(subDT, n) )
```










## Reported weekly events

```{r}
# Snapshot of the last date entered for each country
subDT <- merge(world_map, 
               natcasedeath[year_week == endweek, .(country, dateRep, cases_weekly, deaths_weekly, popData2019)], 
               by.x='region', by.y='countriesAndTerritories', all.x=TRUE, all.y=FALSE)



casedist[countriesAndTerritories == 'Austria' & year_week==endweek,]
natcasedeath[country == 'Austria' & year_week==endweek,]

```

Per `r P/1e6`M residents.

### Global

```{r, fig.width=10, fig.height=10}
p1 <- ggplot(subDT[Date==current, ], aes(x=long, y=lat, group=group, fill=norm_roll_Cases)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=case_col, low='white') +
	coord_cartesian(ylim=c(-60,80)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

p2 <- ggplot(subDT[Date==current, ], aes(x=long, y=lat, group=group, fill=norm_roll_Deaths)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=death_col, low='white') +
	coord_cartesian(ylim=c(-60,80)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

print( p1 / p2 )

ggsave('~/covid/global_roll.png', plot = last_plot(), scale = 1,
       width = 30, height = 30, units = "cm", dpi = 300, limitsize = TRUE)
```

```{r, fig.width=7, fig.height=11}
p1 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Cases)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=case_col, low='white') +
  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

p2 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Deaths)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=death_col, low='white') +
  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

print( p1 / p2 )

ggsave('~/covid/eu_roll.png', plot = last_plot(), scale = 1,
       width = 21, height = 36, units = "cm", dpi = 300, limitsize = TRUE)
```

```{r, fig.width=7, fig.height=6}
# Testing data seems to lag in date yet a bit more than case data.
subDT <- merge(world_df, DT[Date==current-1, ], by.x='id', by.y='countryCode', all=TRUE)
setorder(subDT, ord)

if(any(is.finite(subDT$norm_roll_Tests) & subDT$norm_roll_Tests > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Tests)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=test_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )
  
  ggsave('~/covid/ertest.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}
```

```{r, fig.width=7, fig.height=6}
if(any(!is.na(subDT$norm_roll_Hosp) & subDT$norm_roll_Hosp > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Hosp)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=hosp_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )

    ggsave('~/covid/erhosp.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}

if(any(!is.na(subDT$norm_roll_ICU) & subDT$norm_roll_ICU > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_ICU)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=icu_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )
  
  ggsave('~/covid/ericu.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}
```



## Countries of personal interest

```{r}
setorder(DT, Country, Date)
```

Normalized to `r P` residents.

### In last `r W` days

```{r}
DT[Date==current & Country %in% topInterest, .(Country, norm_roll_Cases, norm_roll_Deaths, norm_roll_Tests, norm_roll_Hosp, norm_roll_ICU)]
```

### Since the beginning

```{r}
DT[Date==current & Country %in% topInterest, .(Country, norm_tot_Cases, norm_tot_Deaths, norm_tot_Tests, norm_tot_Hosp, norm_tot_ICU)]
```

# Timeline

```{r}
# Numbers over time
tidyDT <- melt(DT[, .(Date, Country, 
                      norm_new_Cases, norm_new_Deaths, norm_new_Tests, #norm_new_Hosp, norm_new_ICU,
                      norm_tot_Cases, norm_tot_Deaths, norm_tot_Tests, #norm_tot_Hosp, norm_tot_ICU,
                      norm_roll_Cases, norm_roll_Deaths, norm_roll_Tests#, norm_roll_Hosp, norm_roll_ICU
                      )],
								id.vars = c('Date', 'Country'), variable.name = 'Type', value.name = 'Normalized_count')
tidyDT[grepl('Death', Type), vsplit := 'Deaths']
tidyDT[grepl('Cases', Type), vsplit := 'Cases']
tidyDT[grepl('Tests', Type), vsplit := 'Tests']
tidyDT[, hsplit := sub('_Cases|_Deaths|_Tests|_Hosp|_ICU', '', sub('norm_', '', Type), perl=TRUE)]
setkey(tidyDT, Country)

# Rolling VS Total (exponential-ness)
expDT <- DT[, .(Date, Country, norm_tot_Cases, norm_tot_Deaths, norm_roll_Cases, norm_roll_Deaths)]
tmptot <- melt(expDT, id.vars = c('Date', 'Country'), measure.vars = c('norm_tot_Cases', 'norm_tot_Deaths'), variable.name = 'Type', value.name = 'norm_Total')
tmptot[grepl('Death', Type), vsplit := 'Deaths']
tmptot[!grepl('Death', Type), vsplit := 'Cases']
tmproll <- melt(expDT, id.vars = c('Date', 'Country'), measure.vars = c('norm_roll_Cases', 'norm_roll_Deaths'), variable.name = 'Type', value.name = 'norm_Roll')
tmproll[grepl('Death', Type), vsplit := 'Deaths']
tmproll[!grepl('Death', Type), vsplit := 'Cases']
expDT <- merge(tmproll, tmptot, by=c('Date','Country', 'vsplit'))
setkey(expDT, Country)

# Rate of change
rateDT <- DT[, .(Date, Country, roll_Cases_Rate, roll_Deaths_Rate)]
rateDT <- melt(rateDT, id.vars = c('Date', 'Country'), value.name = 'daily_Rate', variable.name = 'Type')
rateDT[grepl('Death', Type), vsplit := 'Deaths']
rateDT[!grepl('Death', Type), vsplit := 'Cases']
setkey(rateDT, Country)

# Singular plots by information type.
relative_plot <- function(df=expDT, sel_country, title='') {
  ggplot(df, aes(x=norm_Total, y=norm_Roll, group=Country, colour=vsplit)) +
	  facet_grid(vsplit ~ ., scales='free_y') +
		geom_line(colour='black', alpha=0.1, size=0.2) +
		geom_line(data=df[sel_country,], size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
		scale_x_log10() +
		scale_y_log10() +
		coord_cartesian(xlim=c(1,NA), ylim=c(1,NA)) +
		annotation_logticks(sides='lrb') +
		labs(title=title) +
		theme_bw() +
    theme(panel.grid=element_blank(),
          legend.position = 'none')
}

numbers_plot <- function(df=tidyDT, sel_country, title='', linear=FALSE) {
  p1 <- ggplot(df, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
		facet_grid(vsplit ~ hsplit, scales = 'free_y') +
		# geom_line(colour='#000000', alpha=0.1, size=0.2) +  ## Plotting all countries as background for 9 facets
                                                          ## demands too much memory and crashes.
		geom_line(data=df[sel_country,], size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    # coord_cartesian(ylim=c(1,NA)) +
		labs(title=title, y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					panel.grid.minor = element_blank())
  
  if (linear) {
    p1
  } else {
    p1 + scale_y_log10() +
      annotation_logticks(sides='lr')
  }
}

rate_plot <- function(df=rateDT, sel_country, title='', D=W) {
	ggplot(df, aes(x=Date, y=daily_Rate, group=Country, colour=vsplit)) +
    facet_grid(vsplit ~ ., scales='free_y') +
		geom_hline(yintercept = 1, size=0.2) +
	  geom_point(data=df[sel_country,], size=0.5) +
		geom_smooth(data=df[sel_country,], span=0.3, size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(0.6, 1.5)) +
		labs(title=title, x='', y='change_Ratio') +
		theme_bw() +
    theme(legend.position = 'none')
}

ratio_plot <- function(df=tidyDT, sel_country, ref_country=homecountry, title='') {
  ratioDT <- merge(df[sel_country], df[ref_country,], by=c('Date', 'Type'))
  ratioDT[, relative := Normalized_count.x / Normalized_count.y]
  ggplot(ratioDT[hsplit.y=='roll',], aes(x=Date, y=relative, colour=vsplit.y)) +
		facet_grid(vsplit.y ~ hsplit.y, scales = 'free_y') +
		geom_hline(yintercept = 1, size=0.15) +
		geom_point(size=0.5) +
	  geom_smooth(span=0.3, size=0.75) +
		scale_y_continuous(trans='log2', breaks=c(1/64, 1/16, 1/4, 1, 4, 16, 64)) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
		coord_cartesian(ylim=c(1/33, 33)) +
		# annotation_logticks(sides='lr') +
		labs(title=title, y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					axis.text.y.left = element_text())
}

## TODO : Add Hosp/Test, ICU/Hosp, Death/ICU rates to more_plots()

# Collections of singular plots.
more_plots <- function(numbers=tidyDT, exponential=expDT, rates=rateDT, sel_country, ref_country=homecountry){

  cat(sel_country)
  p1 <- relative_plot(exponential, sel_country, 
                      title=paste(sel_country, ': ', W, '-day rolling sum, per', P/1e6, 'M residents, as function of total'))
	p2 <- numbers_plot(numbers[,], sel_country, 
	                   title=paste(sel_country, ': new, total &', W, '-day rolling sum, normalized to', P/1e6, 'M residents'))
	p3 <- rate_plot(rates, sel_country,
	                title=paste(sel_country, ': Daily change rate of', W, '-day roll. sum'))
	p4 <- ratio_plot(numbers, sel_country, ref_country,
	                 title=paste(sel_country,':', W, '-day norm.roll. sum, relative to', ref_country))
	
	print( p1 / p2 / p3 / p4 )
}

fewer_plots <- function(numbers=tidyDT[hsplit=='roll',], exponential=expDT, rates=rateDT, sel_country, ref_country=homecountry){

  cat(sel_country)
  p1 <- relative_plot(exponential, sel_country,
                      title=paste(sel_country, ': ', W, '-day rolling sum, per', P/1e6, 'M residents, as function of total'))
	p2 <- numbers_plot(numbers, sel_country, linear=TRUE,
	                   title=paste(sel_country, ': new, total &', W, '-day rolling sum, normalized to', P/1e6, 'M residents'))
	# p3 <- rate_plot(rates, sel_country,
	                # title=paste(sel_country, ': Daily change rate of', W, '-day roll. sum'))
  
	p4 <- ratio_plot(numbers, sel_country, ref_country, 
	                 title=paste(sel_country,':', W, '-day norm.roll. sum, relative to', ref_country))
  
	print( p1 / p2 / p4)
}

# Easier comparison plots. Many countries by info type.

compare_numbers <- function(df=tidyDT, info, countries) {
  subdf <- df[Type %in% info & Country %in% countries, ]
  m1 <- max(subdf[Date==current, Normalized_count], na.rm = TRUE)
  
  p1 <- ggplot(subdf, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
    geom_hline(yintercept=m1, size=0.25, colour='black', linetype='dotted') +
    geom_line() +
    geom_point(data=subdf[Normalized_count==m1 & Date==current,], colour='black') +
    facet_grid(Country ~ ., ) +
    scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    theme_bw() +
    theme(legend.position = 'none')
  
  p2 <- p1 +   
    scale_y_log10() +
    annotation_logticks(side='lr') +
    coord_cartesian(ylim=c(1,NA))
  
  print((p1 + labs(y=paste(info, '(linear scale)'))) + 
        (p2 + labs(y=paste(info, '(log scale)'))))  
}

compare_exp <- function(df, type, info_x, info_y, countries) {
  subdf <- df[vsplit %in% type & Country %in% countries, ]
  m1 <- max(subdf[Date==current, c(info_y), with=FALSE], na.rm = TRUE)
  m2 <- max(subdf[Date==current, c(info_x), with=FALSE], na.rm = TRUE)
  
  p1 <- ggplot(subdf, aes_string(x=info_x, y=info_y, group='Country', colour='vsplit')) +
    geom_hline(yintercept=m1, size=0.25, colour='black', linetype='dotted') +
    geom_vline(xintercept=m2, size=0.25, colour='black', linetype='dotted') +
    geom_line() +
    geom_point(data=subdf[(subdf[[info_x]]==m2 | subdf[[info_y]]==m1) & Date==current,], colour='black') +
    facet_grid(Country ~ ., ) +
    scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    theme_bw() +
    theme(legend.position = 'none')
  
  p2 <- p1 +   
    scale_y_log10() +
    scale_x_log10() +
    annotation_logticks(side='lb') +
    coord_cartesian(ylim=c(1,NA), xlim=c(1,NA))
  
  print((p1 + labs(y=paste(info_y, type, '(linear scale)'), x=paste(info_x, type, '(linear scale)') )) + 
        (p2 + labs(y=paste(info_y, type, '(log scale)'), x=paste(info_x, type, '(linear scale)') )))  
}

```

## Global

```{r, fig.height=8, fig.width=8, warning=FALSE}
subDT <- melt(unique(DT[, .(Date, gNew_Cases, gNew_Deaths, gTotal_Cases, gTotal_Deaths, gRoll_Cases, gRoll_Deaths)]),
							id.vars="Date", variable.name="Type", value.name="Events")

p1 <- ggplot(subDT, aes(x=Date, y=Events, colour=Type, fill=Type)) +
	facet_grid( sub('^g', '', sub('_Cases', '', sub('_Deaths', '', subDT$Type))) ~ ., scales = 'free_y') +
	geom_line() +
	theme_minimal() + 
	labs(x='', y='') +
	theme(legend.position='none')

p2 <- ggplot(subDT, aes(x=Date, y=Events, colour=Type, fill=Type)) +
	facet_grid( sub('^g', '', sub('_Cases', '', sub('_Deaths', '', subDT$Type))) ~ ., scales = 'free_y') +
	geom_line() +
	scale_y_log10() +
	labs(x='', y='') +
	theme_minimal()

print( p1 + p2 )

subDT <- unique(DT[, .(Date, gRoll_Cases_Rate, gRoll_Deaths_Rate, gMortality)])

p3 <- ggplot(subDT, aes(x=Date, y=gRoll_Cases_Rate)) +
	geom_hline(yintercept = 1, size=0.1) +
  geom_point(colour=case_col, size=0.5) +
	geom_smooth(span=0.2, colour=case_col, size=0.5) +
	coord_cartesian(ylim=c(0.8, 1.2)) +
	labs(title=paste('Daily change rate of ', W, '-day rolling sum')) +
	theme_bw()

p4 <- ggplot(subDT, aes(x=Date, y=gRoll_Deaths_Rate)) +
	geom_hline(yintercept = 1, size=0.1) +
  geom_point(colour=death_col, size=0.5) +
	geom_smooth(span=0.2, colour=death_col, size=0.5) +
	coord_cartesian(ylim=c(0.8, 1.2)) +
	labs(title='') +
		theme_bw()

p5 <- ggplot(subDT, aes(x=Date, y=gMortality)) +
	geom_hline(yintercept = 0, size=0.1) +
  geom_point(colour=death_col, size=0.5) +
	geom_smooth(span=0.3, colour=death_col, size=0.5) +
	labs(title='') +
		theme_bw()

print( p3 / p4 / p5 )
```

## Detailed

### Countries of personal interest

#### More plots

```{r, fig.height=20, fig.width=10, warning=FALSE}
for (i in topInterest) {
	more_plots(tidyDT, expDT, rateDT, i, homecountry)
  
  ggsave(paste0('~/covid/', i, '_more.png'), plot = last_plot(), scale = 1,
       width = 20, height = 40, units = "cm", dpi = 300, limitsize = TRUE)
}
```

#### Fewer plots

```{r, fig.height=15, fig.width=10, warning=FALSE}
for (i in topInterest) {
	fewer_plots(tidyDT[hsplit=='roll',], expDT, rateDT, i, homecountry)
  
  ggsave(paste0('~/covid/', i, '_fewer.png'), plot = last_plot(), scale = 1,
       width = 20, height = 40, units = "cm", dpi = 300, limitsize = TRUE)
}
```

## Comparative

### Selected countries

```{r, fig.height=20, fig.width=10, warning=FALSE}
countries <- c(topInterest, other_countries, neighbour_countries)

for (j in c('norm_roll_Cases', 'norm_roll_Deaths', 'norm_new_Tests')) {
	compare_numbers(tidyDT, j, countries)
  
  ggsave(paste0('~/covid/', j, '_comp.png'), plot = last_plot(), scale = 1,
       width = 30, height = 3 * length(countries), units = "cm", dpi = 300, limitsize = TRUE)
}

for (j in c('Cases', 'Deaths')) {
  compare_exp(expDT, j, 'norm_Total', 'norm_Roll', countries)
    
  ggsave(paste0('~/covid/', j, '_comp.png'), plot = last_plot(), scale = 1,
       width = 30, height = 3 * length(countries), units = "cm", dpi = 300, limitsize = TRUE)
  }
```


# Session

```{r}
sessionInfo()
```
