---
title: "COVID19"
author: "Kimon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

# Libraries

```{r message=FALSE}
library(httr)
library(lubridate)
library(ggplot2)
library(gganimate)
library(ggrepel)
library(patchwork)
library(data.table)
library(broom)
library(rgdal)
require(maptools)
require(rgeos)

options(scipen=2)
```

Source: https://opendata.ecdc.europa.eu/covid19/

# Load and process map and current data

```{r}
tf <- "~/covid/global.csv"
GET("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", authenticate(":", ":", type="ntlm"), write_disk(tf, overwrite = TRUE))
globcases <- fread(tf)
setnames(globcases, c('dateRep', 'day', 'month', 'year', 'new_Cases', 'new_Deaths', 'Country', 'geoId', 'countryCode', 'population', 'continent', 'roll_norm_Cases'))
globcases[, Date := as.Date(dmy(dateRep))]
globcases <- globcases[, .(Country, countryCode, continent, population, Date, new_Cases, new_Deaths)]
# str(globcases)

## Create year-week field, to enable smoother merging with the other tables, as they can have NA dates.
# Start with separate fields, in case I need to add tweaks to them.
globcases[, year := strftime(globcases$Date, format='%Y')]
globcases[, week := as.integer(strftime(globcases$Date, format='%V'))]
globcases[, year_week := paste0(year, '-W', week)]
globcases[, year := NULL]
globcases[, week := NULL]

# Remove the cruise ship entry
globcases <- globcases[Country != 'Cases_on_an_international_conveyance_Japan', ]

# Some countries lag by a day in entering their data. So if I use the latest date I miss some countries,
# if I use the latest available entry, the countries are out of sync. Determine range, and as long as it's less than a couple days, take the latest common date as the "current" date.
end_dates <- globcases[, max(Date), by=Country]

if ( min(unique(end_dates$V1)) - today() >= -3) { # allow for lack of updates over weekends
  message('Entries for some countries lag by a couple days. Continuing with last common date.')
  current <- min(unique(end_dates$V1))
} else {
  message('Entries for some countries seem to have stopped. Continuing with latest available entry date.')
  current <- max(unique(end_dates$V1))
}

message(paste('Last entry date used:', current, '.'))


tf <- "~/covid/testing.csv"
GET("https://opendata.ecdc.europa.eu/covid19/testing/csv", authenticate(":", ":", type="ntlm"), write_disk(tf, overwrite = TRUE))
tests <- fread(tf)
setnames(tests, c('Country', 'countryCode', 'year_week', 'test_Cases', 'new_Tests', 'population', 'test_Rate', 'pos_Rate', 'source'))
tests <- tests[, .(Country, year_week, new_Tests)]
# str(tests)
## The UK has been mistyped in some entries, so that needs fixing to enable better correct merging of the tables.
tests[Country=='United Kingdom', Country := 'United_Kingdom']
## Fix inconsistent formatting that is messing up the mergers
tests[, year_week:= sub('W0', 'W', year_week)]

tf <- "~/covid/hospitalizations.csv"
GET("https://opendata.ecdc.europa.eu/covid19/hospitalicuadmissionrates/csv", authenticate(":", ":", type="ntlm"), write_disk(tf, overwrite = TRUE))
icu <- fread(tf)
setnames(icu, c('Country', 'Type', 'Date', 'year_week', 'Value', 'source', 'URL'))
## The UK has been mistyped in some entries, so that needs fixing to enable better correct merging of the tables.
icu[Country=='United Kingdom', Country := 'United_Kingdom']
icu[, Date := as.Date(Date)]
## Fix inconsistent formatting that is messing up the mergers
icu[, year_week:= sub('W0', 'W', year_week)]

## Split daily and weekly metrics that are wreaking havoc on my mergers.
occup <- unique(icu[grepl('occupancy', Type, fixed=TRUE), .(Country, Date, Type, Value)])
occup <- dcast(occup, Country + Date ~ Type, value.var = 'Value')
setnames(occup, c('Country', 'Date', 'occup_ICU', 'occup_Hosp'))
admit <- unique(icu[grepl('admissions', Type, fixed=TRUE), .(Country, year_week, Type, Value)])
admit <- dcast(admit, Country + year_week ~ Type, value.var = 'Value')
setnames(admit, c('Country', 'year_week', 'norm_new_ICU', 'norm_new_Hosp'))
admit[, norm_new_ICU := norm_new_ICU * 10 ]   # re-normalize to 1M instead of 100K, to be same as my other norms
admit[, norm_new_Hosp := norm_new_Hosp * 10 ]
# str(icu)
```

```{r}
# Marry the tables
DT <- merge(merge(globcases, occup, by=c('Country', 'Date'), all=TRUE), 
            merge(admit, tests, by=c('Country', 'year_week'), all=TRUE), 
            by=c('Country', 'year_week'), all=TRUE)
# View(merge(globcases, occup, by=c('Country', 'Date'), all=TRUE)[Country=='Denmark',])
# View(merge(admit, tests, by=c('Country', 'year_week'), all=TRUE)[Country=='Denmark',])
# View(DT[Country=='Denmark',])
# str(DT)

## Some country codes are NA in some rows. Try to fill in the missing values.
DT[, countryCode := .SD$countryCode[!is.na(.SD$countryCode)][1], by=Country]

# Sort newest last, for cumulative metrics.
setorder(DT, Country, Date)

## Missing days break the cumulative sums
DT[is.na(new_Cases), new_Cases := 0]
DT[is.na(new_Deaths), new_Deaths := 0]
DT[is.na(new_Tests), new_Tests := 0]
DT[is.na(norm_new_Hosp), norm_new_Hosp := 0]
DT[is.na(norm_new_ICU), norm_new_ICU := 0]

# Total Cases and Deaths
DT[, total_Cases := cumsum(new_Cases), by=Country]
DT[, total_Deaths := cumsum(new_Deaths), by=Country]
<<<<<<< HEAD

# Sliding window cumulative cases in a W-days window.
W <- 14
DT[, roll_Cases := frollsum(new_Cases, n=W, align='right'), by=Country]
DT[, roll_Deaths := frollsum(new_Deaths, n=W, align='right'), by=Country]

# Day relative to first reported case
DT[, Day := 1:length(.SD$new_Cases), by=Country]

# Day relative to N deaths
N <- 100
DT[, Nplus := abs(total_Deaths - N)]
DT[, Day_Aligned := Day - .SD[Nplus==min(Nplus), Day][1], by=Country]
DT[, Nplus := NULL]

# Day relative to today
DT[, past_Days := Day - max(Day), by=Country]
=======
DT[, total_Tests := cumsum(new_Tests), by=Country]

# Sliding window cumulative cases in a W-days window.
W <- 7
DT[, roll_Cases := frollsum(new_Cases, n=W, align='right', na.rm=TRUE), by=Country]
DT[, roll_Deaths := frollsum(new_Deaths, n=W, align='right', na.rm=TRUE), by=Country]
DT[, roll_Tests := frollsum(new_Tests, n=W, align='right', na.rm=TRUE), by=Country]
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8

# Population normalisations to P citizens
P <- 1e6
DT[, norm_new_Cases := new_Cases / population * P, by=Country]
DT[, norm_new_Deaths := new_Deaths / population * P, by=Country]
DT[, norm_new_Tests := new_Tests / population * P, by=Country]
# DT[is.na(norm_new_Cases), norm_new_Cases := 0]
# DT[is.na(norm_new_Deaths), norm_new_Deaths := 0]
# DT[is.na(norm_new_Tests), norm_new_Tests := 0]

DT[, norm_tot_Cases := total_Cases / population * P, by=Country]
DT[, norm_tot_Deaths := total_Deaths / population * P, by=Country]
DT[, norm_tot_Tests := total_Tests / population * P, by=Country]
DT[, norm_tot_ICU := cumsum(norm_new_ICU), by=Country]
DT[, norm_tot_Hosp := cumsum(norm_new_Hosp), by=Country]
# DT[is.na(norm_tot_Cases), norm_tot_Cases := 0]
# DT[is.na(norm_tot_Deaths), norm_tot_Deaths := 0]
# DT[is.na(norm_tot_Tests), norm_tot_Tests := 0]
# DT[is.na(norm_tot_ICU), norm_tot_ICU := 0]
# DT[is.na(norm_tot_Hosp), norm_tot_Hosp := 0]

DT[, norm_roll_Cases := roll_Cases / population * P, by=Country]
DT[, norm_roll_Deaths := roll_Deaths / population * P, by=Country]
DT[, norm_roll_Tests := roll_Tests / population * P, by=Country]
DT[, norm_roll_ICU := frollsum(norm_new_ICU, n=W, align='right', na.rm=TRUE), by=Country]
DT[, norm_roll_Hosp := frollsum(norm_new_Hosp, n=W, align='right', na.rm=TRUE), by=Country]

# Severity rates
DT[, Mortality := total_Deaths / total_Cases]

DT[, roll_TestCase_Rate := norm_roll_Cases / norm_roll_Tests]
DT[, roll_HospICU_Rate := norm_roll_ICU / norm_roll_Hosp]
DT[, roll_ICUDeath_Rate := norm_roll_Deaths / norm_roll_ICU]

# Global
DT[, gNew_Cases := sum(new_Cases, na.rm=TRUE), by=Date]
DT[, gNew_Deaths := sum(new_Deaths, na.rm=TRUE), by=Date]
DT[, gNew_Tests := sum(new_Tests, na.rm=TRUE), by=Date]
DT[, gTotal_Cases := sum(total_Cases, na.rm=TRUE), by=Date]
DT[, gTotal_Deaths := sum(total_Deaths, na.rm=TRUE), by=Date]
DT[, gTotal_Tests := sum(total_Tests, na.rm=TRUE), by=Date]
DT[, gRoll_Cases := sum(roll_Cases, na.rm=TRUE), by=Date]
DT[, gRoll_Deaths := sum(roll_Deaths, na.rm=TRUE), by=Date]
DT[, gRoll_Tests := sum(roll_Tests, na.rm=TRUE), by=Date]
DT[, gMortality := gTotal_Deaths / gTotal_Cases]

# Rates of daily change
DT[, roll_Cases_Rate := frollapply(roll_Cases, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[, roll_Deaths_Rate := frollapply(roll_Deaths, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[, roll_Tests_Rate := frollapply(norm_roll_Tests, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[, roll_Hosp_Rate := frollapply(norm_roll_Hosp, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[, roll_ICU_Rate := frollapply(norm_roll_ICU, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[!is.finite(roll_Cases_Rate), roll_Cases_Rate := 0]
DT[!is.finite(roll_Deaths_Rate), roll_Deaths_Rate := 0]
DT[!is.finite(roll_Tests_Rate), roll_Tests_Rate := 0]
DT[!is.finite(roll_Hosp_Rate), roll_Hosp_Rate := 0]
DT[!is.finite(roll_ICU_Rate), roll_ICU_Rate := 0]

DT[, gRoll_Cases_Rate := frollapply(gRoll_Cases, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[, gRoll_Deaths_Rate := frollapply(gRoll_Deaths, n=2, FUN = function(x){x[2] / x[1]}, align="right"), by=Country]
DT[!is.finite(gRoll_Cases_Rate), gRoll_Cases_Rate := 0]
DT[!is.finite(gRoll_Deaths_Rate), gRoll_Deaths_Rate := 0]

# Compare severity to one country.
homecountry = 'Austria'
<<<<<<< HEAD
=======
# Select countries
topInterest <- c('Austria', 'Italy', 'Greece', 'Luxembourg', 'Germany')
neighbour_countries <- c('Switzerland', 'Slovakia', 'Slovenia', 'Czechia', 'Hungary')
other_countries <- c('United_Kingdom', 'Spain', 'France', 'Belgium', 'United_States_of_America', 'Sweden', 'South_Korea')
```

```{r}
# Download the shape file from the web and unzip it:
# download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip", destfile="~/covid/shape_files/world_shape_file.zip")
# system("unzip ~/covid/shape_files/world_shape_file.zip")
world_spdf <- readOGR(dsn='~/covid/shape_files/world_shape_file', layer='TM_WORLD_BORDERS_SIMPL-0.3')

# Dataframe-ize map.
world_df <- as.data.table(tidy(world_spdf, region="ISO3"))
# For plottting, the order of rows is super important. 
# Merge operations later can change the order, so I need to be able to recover it.
world_df[, ord := 1:nrow(world_df)]
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
```

The Covid19 dataset from ECDC comes without geospacial data. The geospacial data available from other sources may not be the most up-to-date with recognised countries and names.

```{r}
message(paste(sum(!(world_spdf$ISO3 %in% DT$countryCode)), 
							"countries in the map file do not correspond to an entry in the Covid19 data."))

print(world_spdf$ISO3[! world_spdf$ISO3 %in% DT$countryCode])


outgroup <- unique(DT[!(countryCode %in% world_spdf$ISO3), .(Country, countryCode)])
message(paste(nrow(outgroup),
							"countries in the Covid19 data do not correspond to an entity in the map file:"))
print(outgroup)
```

I think for a global overview map, the missing countries will not make a big difference.


# Status

```{r}
<<<<<<< HEAD
minpop=5e5
=======
case_col = '#FF0000'
death_col = '#0055FF'
hosp_col = '#555500'
icu_col = '#995500'
test_col = '#005500'
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8

print( data.frame(DaysTracked = length(unique(DT$Date)),
					CountriesTracked = length(unique(DT$Country)) ) )

print( data.frame( Global_Cases = sum(DT$new_Cases, na.rm = TRUE),
					Global_Deaths = sum(DT$new_Deaths, na.rm = TRUE),
					Global_Mortality = sum(DT$new_Deaths, na.rm = TRUE) / sum(DT$new_Cases, na.rm = TRUE) ) )
```

## World worst and best

```{r}
subDT <- DT[Date==current, , by=Country][, .(Country, norm_roll_Cases)]
# Worst
tail(subDT, n=10)
# Best
head(subDT, n=10)

subDT <- DT[Date==current, , by=Country][, norm_roll_Deaths]
# Worst
tail(subDT, n=10)
# Best
head(subDT, n=10)

```

## Europe worst and best

```{r}
subDT <- DT[continent=='Europe', max(norm_roll_Cases, na.rm=TRUE), by=Country][order(V1),]
setnames(subDT, c('Country', 'norm_roll_Cases'))
# Worst
tail(subDT, n=10)
# Best
head(subDT, n=10)
```


## Reported events in last `r W` days

```{r}
case_col = '#FF0000'
death_col = '#0055FF'
hosp_col = '#555500'
icu_col = '#995500'
test_col = '#005500'

subDT <- merge(world_df, DT[Date==current, ], by.x='id', by.y='countryCode', all=TRUE)
setorder(subDT, ord)
```

Per `r P/1e6`M residents.

```{r, fig.width=10, fig.height=10}
p1 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Cases)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=case_col, low='white') +
	coord_cartesian(ylim=c(-60,80)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

p2 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Deaths)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=death_col, low='white') +
	coord_cartesian(ylim=c(-60,80)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

print( p1 / p2 )

ggsave('~/covid/global_roll.png', plot = last_plot(), scale = 1,
       width = 30, height = 30, units = "cm", dpi = 300, limitsize = TRUE)
```

```{r, fig.width=7, fig.height=11}
p1 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Cases)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=case_col, low='white') +
  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

p2 <- ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Deaths)) +
	geom_polygon(colour='black', size=0.2) +
	scale_fill_gradient(high=death_col, low='white') +
  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
	theme_void() +
	theme(panel.background = element_rect(fill='#BBFFFF'))

print( p1 / p2 )

ggsave('~/covid/eu_roll.png', plot = last_plot(), scale = 1,
       width = 21, height = 36, units = "cm", dpi = 300, limitsize = TRUE)
```

<<<<<<< HEAD
## Countries of personal interest

```{r}
topInterest <- c('Austria', 'Italy', 'Greece', 'Germany', 'Luxembourg')
setorder(DT, Country, past_Days)
=======
```{r, fig.width=7, fig.height=6}
# Testing data seems to lag in date yet a bit more than case data.
subDT <- merge(world_df, DT[Date==current-1, ], by.x='id', by.y='countryCode', all=TRUE)
setorder(subDT, ord)

if(any(is.finite(subDT$norm_roll_Tests) & subDT$norm_roll_Tests > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Tests)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=test_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )
  
  ggsave('~/covid/ertest.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}
```

```{r, fig.width=7, fig.height=6}
if(any(!is.na(subDT$norm_roll_Hosp) & subDT$norm_roll_Hosp > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_Hosp)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=hosp_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )

    ggsave('~/covid/erhosp.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}

if(any(!is.na(subDT$norm_roll_ICU) & subDT$norm_roll_ICU > 0)){
  print(
    ggplot(subDT, aes(x=long, y=lat, group=group, fill=norm_roll_ICU)) +
  		geom_polygon(colour='black', size=0.2) +
  		scale_fill_gradient(high=icu_col, low='white') +
  	  coord_cartesian(xlim=c(-10, 50), ylim=c(30,70)) +
  		theme_void() +
  		theme(panel.background = element_rect(fill='#BBFFFF'))
  )
  
  ggsave('~/covid/ericu.png', plot = last_plot(), scale = 1,
       width = 21, height = 18, units = "cm", dpi = 300, limitsize = TRUE)
}
```



## Countries of personal interest

```{r}
setorder(DT, Country, Date)
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
```

Normalized to `r P` residents.

### In last `r W` days

```{r}
DT[Date==current & Country %in% topInterest, .(Country, norm_roll_Cases, norm_roll_Deaths, norm_roll_Tests, norm_roll_Hosp, norm_roll_ICU)]
```

### Since the beginning

```{r}
DT[Date==current & Country %in% topInterest, .(Country, norm_tot_Cases, norm_tot_Deaths, norm_tot_Tests, norm_tot_Hosp, norm_tot_ICU)]
```

# Timeline

```{r}
## TODO: Add ICU/Hosp to tidyDT

# Numbers over time
tidyDT <- melt(DT[, .(Date, Country, 
                      norm_new_Cases, norm_new_Deaths, norm_new_Tests, #norm_new_Hosp, norm_new_ICU,
                      norm_tot_Cases, norm_tot_Deaths, norm_tot_Tests, #norm_tot_Hosp, norm_tot_ICU,
                      norm_roll_Cases, norm_roll_Deaths, norm_roll_Tests#, norm_roll_Hosp, norm_roll_ICU
                      )],
								id.vars = c('Date', 'Country'), variable.name = 'Type', value.name = 'Normalized_count')
tidyDT[grepl('Death', Type), vsplit := 'Deaths']
tidyDT[grepl('Cases', Type), vsplit := 'Cases']
tidyDT[grepl('Tests', Type), vsplit := 'Tests']
# tidyDT[grepl('Hosp', Type), vsplit := 'Hosp']
# tidyDT[grepl('ICU', Type), vsplit := 'ICU']
tidyDT[, hsplit := sub('_Cases|_Deaths|_Tests|_Hosp|_ICU', '', sub('norm_', '', Type), perl=TRUE)]
setkey(tidyDT, Country)

# Rolling VS Total (exponential-ness)
expDT <- DT[, .(Date, Country, norm_tot_Cases, norm_tot_Deaths, norm_roll_Cases, norm_roll_Deaths)]
tmptot <- melt(expDT, id.vars = c('Date', 'Country'), measure.vars = c('norm_tot_Cases', 'norm_tot_Deaths'), variable.name = 'Type', value.name = 'norm_Total')
tmptot[grepl('Death', Type), vsplit := 'Deaths']
tmptot[!grepl('Death', Type), vsplit := 'Cases']
tmproll <- melt(expDT, id.vars = c('Date', 'Country'), measure.vars = c('norm_roll_Cases', 'norm_roll_Deaths'), variable.name = 'Type', value.name = 'norm_Roll')
tmproll[grepl('Death', Type), vsplit := 'Deaths']
tmproll[!grepl('Death', Type), vsplit := 'Cases']
expDT <- merge(tmproll, tmptot, by=c('Date','Country', 'vsplit'))
setkey(expDT, Country)

# Rate of change
rateDT <- DT[, .(Date, Country, roll_Cases_Rate, roll_Deaths_Rate)]
rateDT <- melt(rateDT, id.vars = c('Date', 'Country'), value.name = 'daily_Rate', variable.name = 'Type')
rateDT[grepl('Death', Type), vsplit := 'Deaths']
rateDT[!grepl('Death', Type), vsplit := 'Cases']
setkey(rateDT, Country)

# Singular plots by information type.
relative_plot <- function(df=expDT, sel_country, title='') {
  ggplot(df, aes(x=norm_Total, y=norm_Roll, group=Country, colour=vsplit)) +
	  facet_grid(vsplit ~ ., scales='free_y') +
		geom_line(colour='black', alpha=0.1, size=0.2) +
		geom_line(data=df[sel_country,], size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
		scale_x_log10() +
		scale_y_log10() +
		coord_cartesian(xlim=c(1,NA), ylim=c(1,NA)) +
		annotation_logticks(sides='lrb') +
		labs(title=title) +
		theme_bw() +
    theme(panel.grid=element_blank(),
          legend.position = 'none')
}

numbers_plot <- function(df=tidyDT, sel_country, title='', linear=FALSE) {
  p1 <- ggplot(df, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
		facet_grid(vsplit ~ hsplit, scales = 'free_y') +
		# geom_line(colour='#000000', alpha=0.1, size=0.2) +  ## Plotting all countries as background for 9 facets
                                                          ## demands too much memory and crashes.
		geom_line(data=df[sel_country,], size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    # coord_cartesian(ylim=c(1,NA)) +
		labs(title=title, y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					panel.grid.minor = element_blank())
  
  if (linear) {
    p1
  } else {
    p1 + scale_y_log10() +
      annotation_logticks(sides='lr')
  }
}

<<<<<<< HEAD
mDT <- melt(DT[, .(Date, Country, norm_new_Cases, norm_new_Deaths, norm_tot_Cases, norm_tot_Deaths, norm_roll_Cases, norm_roll_Deaths)],
							id.vars = c('Date', 'Country'), variable.name = 'Type', value.name = 'Normalized_count')

# Countries enter the table at different dates. That is inconvenient.
# Fix by unmelting Countries so that missing bits get padded and then re-melting
# mDT <- melt(dcast(mDT, Date ~ Country, value.var="Normalized_count", fill=NA),
# 		 id.vars='Date', variable.name='Country', value.name='Normalized_count')

mDT[grepl('Death', Type), vsplit := 'Deaths']
mDT[!grepl('Death', Type), vsplit := 'Cases']
mDT[, hsplit := sub('_Cases|_Deaths', '', sub('norm_', '', Type), perl=TRUE)]

=======
rate_plot <- function(df=rateDT, sel_country, title='', D=W) {
	ggplot(df, aes(x=Date, y=daily_Rate, group=Country, colour=vsplit)) +
    facet_grid(vsplit ~ ., scales='free_y') +
		geom_hline(yintercept = 1, size=0.2) +
	  geom_point(data=df[sel_country,], size=0.5) +
		geom_smooth(data=df[sel_country,], span=0.3, size=0.8) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(0.6, 1.5)) +
		labs(title=title, x='', y='change_Ratio') +
		theme_bw() +
    theme(legend.position = 'none')
}

ratio_plot <- function(df=tidyDT, sel_country, ref_country=homecountry, title='') {
  ratioDT <- merge(df[sel_country], df[ref_country,], by=c('Date', 'Type'))
  ratioDT[, relative := Normalized_count.x / Normalized_count.y]
  ggplot(ratioDT[hsplit.y=='roll',], aes(x=Date, y=relative, colour=vsplit.y)) +
		facet_grid(vsplit.y ~ hsplit.y, scales = 'free_y') +
		geom_hline(yintercept = 1, size=0.15) +
		geom_point(size=0.5) +
	  geom_smooth(span=0.3, size=0.75) +
		scale_y_continuous(trans='log2', breaks=c(1/64, 1/16, 1/4, 1, 4, 16, 64)) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
		coord_cartesian(ylim=c(1/33, 33)) +
		# annotation_logticks(sides='lr') +
		labs(title=title, y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					axis.text.y.left = element_text())
}

## TODO : Add Hosp/Test, ICU/Hosp, Death/ICU rates to more_plots()

# Collections of singular plots.
more_plots <- function(numbers=tidyDT, exponential=expDT, rates=rateDT, sel_country, ref_country=homecountry){

  cat(sel_country)
  p1 <- relative_plot(exponential, sel_country, 
                      title=paste(sel_country, ': ', W, '-day rolling sum, per', P/1e6, 'M residents, as function of total'))
	p2 <- numbers_plot(numbers[,], sel_country, 
	                   title=paste(sel_country, ': new, total &', W, '-day rolling sum, normalized to', P/1e6, 'M residents'))
	p3 <- rate_plot(rates, sel_country,
	                title=paste(sel_country, ': Daily change rate of', W, '-day roll. sum'))
	p4 <- ratio_plot(numbers, sel_country, ref_country,
	                 title=paste(sel_country,':', W, '-day norm.roll. sum, relative to', ref_country))
	
	print( p1 / p2 / p3 / p4 )
}

fewer_plots <- function(numbers=tidyDT[hsplit=='roll',], exponential=expDT, rates=rateDT, sel_country, ref_country=homecountry){

  cat(sel_country)
  p1 <- relative_plot(exponential, sel_country,
                      title=paste(sel_country, ': ', W, '-day rolling sum, per', P/1e6, 'M residents, as function of total'))
	p2 <- numbers_plot(numbers, sel_country, linear=TRUE,
	                   title=paste(sel_country, ': new, total &', W, '-day rolling sum, normalized to', P/1e6, 'M residents'))
	# p3 <- rate_plot(rates, sel_country,
	                # title=paste(sel_country, ': Daily change rate of', W, '-day roll. sum'))
  
	p4 <- ratio_plot(numbers, sel_country, ref_country, 
	                 title=paste(sel_country,':', W, '-day norm.roll. sum, relative to', ref_country))
  
	print( p1 / p2 / p4)
}

# Easier comparison plots. Many countries by info type.

compare_numbers <- function(df=tidyDT, info, countries) {
  subdf <- df[Type %in% info & Country %in% countries, ]
  m1 <- max(subdf[Date==current, Normalized_count], na.rm = TRUE)
  
  p1 <- ggplot(subdf, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
    geom_hline(yintercept=m1, size=0.25, colour='black', linetype='dotted') +
    geom_line() +
    geom_point(data=subdf[Normalized_count==m1 & Date==current,], colour='black') +
    facet_grid(Country ~ ., ) +
    scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    theme_bw() +
    theme(legend.position = 'none')
  
  p2 <- p1 +   
    scale_y_log10() +
    annotation_logticks(side='lr') +
    coord_cartesian(ylim=c(1,NA))
  
  print((p1 + labs(y=paste(info, '(linear scale)'))) + 
        (p2 + labs(y=paste(info, '(log scale)'))))  
}

compare_exp <- function(df, type, info_x, info_y, countries) {
  subdf <- df[vsplit %in% type & Country %in% countries, ]
  m1 <- max(subdf[Date==current, c(info_y), with=FALSE], na.rm = TRUE)
  m2 <- max(subdf[Date==current, c(info_x), with=FALSE], na.rm = TRUE)
  
  p1 <- ggplot(subdf, aes_string(x=info_x, y=info_y, group='Country', colour='vsplit')) +
    geom_hline(yintercept=m1, size=0.25, colour='black', linetype='dotted') +
    geom_vline(xintercept=m2, size=0.25, colour='black', linetype='dotted') +
    geom_line() +
    geom_point(data=subdf[(subdf[[info_x]]==m2 | subdf[[info_y]]==m1) & Date==current,], colour='black') +
    facet_grid(Country ~ ., ) +
    scale_colour_manual(values=c(Deaths=death_col, Cases=case_col, Tests=test_col, Hosp=hosp_col, ICU=icu_col)) +
    theme_bw() +
    theme(legend.position = 'none')
  
  p2 <- p1 +   
    scale_y_log10() +
    scale_x_log10() +
    annotation_logticks(side='lb') +
    coord_cartesian(ylim=c(1,NA), xlim=c(1,NA))
  
  print((p1 + labs(y=paste(info_y, type, '(linear scale)'), x=paste(info_x, type, '(linear scale)') )) + 
        (p2 + labs(y=paste(info_y, type, '(log scale)'), x=paste(info_x, type, '(linear scale)') )))  
}
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8

```

## Global

```{r, fig.height=8, fig.width=8, warning=FALSE}
subDT <- melt(unique(DT[, .(Date, gNew_Cases, gNew_Deaths, gTotal_Cases, gTotal_Deaths, gRoll_Cases, gRoll_Deaths)]),
							id.vars="Date", variable.name="Type", value.name="Events")

p1 <- ggplot(subDT, aes(x=Date, y=Events, colour=Type, fill=Type)) +
	facet_grid( sub('^g', '', sub('_Cases', '', sub('_Deaths', '', subDT$Type))) ~ ., scales = 'free_y') +
	geom_line() +
	theme_minimal() + 
	labs(x='', y='') +
	theme(legend.position='none')

p2 <- ggplot(subDT, aes(x=Date, y=Events, colour=Type, fill=Type)) +
	facet_grid( sub('^g', '', sub('_Cases', '', sub('_Deaths', '', subDT$Type))) ~ ., scales = 'free_y') +
	geom_line() +
	scale_y_log10() +
	labs(x='', y='') +
	theme_minimal()

print( p1 + p2 )

subDT <- unique(DT[, .(Date, gRoll_Cases_Rate, gRoll_Deaths_Rate, gMortality)])

p3 <- ggplot(subDT, aes(x=Date, y=gRoll_Cases_Rate)) +
	geom_hline(yintercept = 1, size=0.1) +
  geom_point(colour=case_col, size=0.5) +
	geom_smooth(span=0.2, colour=case_col, size=0.5) +
	coord_cartesian(ylim=c(0.8, 1.2)) +
	labs(title=paste('Daily change rate of ', W, '-day rolling sum')) +
	theme_bw()

p4 <- ggplot(subDT, aes(x=Date, y=gRoll_Deaths_Rate)) +
	geom_hline(yintercept = 1, size=0.1) +
  geom_point(colour=death_col, size=0.5) +
	geom_smooth(span=0.2, colour=death_col, size=0.5) +
	coord_cartesian(ylim=c(0.8, 1.2)) +
	labs(title='') +
		theme_bw()

p5 <- ggplot(subDT, aes(x=Date, y=gMortality)) +
	geom_hline(yintercept = 0, size=0.1) +
<<<<<<< HEAD
	geom_line(colour='magenta', size=0.5) +
=======
  geom_point(colour=death_col, size=0.5) +
	geom_smooth(span=0.3, colour=death_col, size=0.5) +
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
	labs(title='') +
		theme_bw()

print( p3 / p4 / p5 )
```

<<<<<<< HEAD
## Home

```{r, fig.height=20, fig.width=10, warning=FALSE}
i <- 'Austria'
cat(i)
	
p1 <- ggplot(mDT, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
	facet_grid(hsplit ~ vsplit, scales = 'free_y') +
	geom_line(colour='#000000', alpha=0.1, size=0.25) +
	geom_line(data=mDT[Country==i,], size=0.75) +
	scale_y_log10() +
	scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
	coord_cartesian(ylim=c(1,NA)) +
	annotation_logticks(sides='lr') +
	labs(title=paste(i, ': new, total &', W, '-day rolling sum, per', P, 'residents'), y='', x='') +
	theme_bw() + 
	theme(legend.position = 'none',
				panel.grid = element_blank())

p2a <- relative_plot(DT, i, 'total_Cases', 'roll_Cases', case_col)
p2b <- relative_plot(DT, i, 'total_Deaths', 'roll_Deaths', death_col)

p3a <- rate_plot(DT, i, 'roll_Cases_Rate', 'Daily change rates', case_col)
p3b <- rate_plot(DT, i, 'roll_Deaths_Rate', '', death_col)

print( p1 / (p2a / p2b) / (p3a + p3b) )
```

```{r}
refDT <- mDT[Country==i,]
# mDT[, nrow(.SD), by=Country]

```

## Countries of personal interest

```{r, fig.height=20, fig.width=10, warning=FALSE}
for (i in topInterest[topInterest != 'Austria']) {
	# i <- topInterest[2]
	cat(i)
	
	p1 <- ggplot(mDT, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
		facet_grid(hsplit ~ vsplit, scales = 'free_y') +
		geom_line(colour='#000000', alpha=0.1, size=0.25) +
		geom_line(data=mDT[Country==i,], size=0.75) +
		scale_y_log10() +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(1,NA)) +
		annotation_logticks(sides='lr') +
		labs(title=paste(i, ': new, total &', W, '-day rolling sum, per', P, 'residents'), y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					panel.grid = element_blank())
	

	p2a <- relative_plot(DT, i, 'total_Cases', 'roll_Cases', case_col)
	p2b <- relative_plot(DT, i, 'total_Deaths', 'roll_Deaths', death_col)

	# p3a <- rate_plot(DT, i, 'roll_Cases_Rate', 'Daily change rates', case_col)
	# p3b <- rate_plot(DT, i, 'roll_Deaths_Rate', '', death_col)

	
	subDT <- merge(refDT, mDT[Country==i,], by=c('Date', 'Type'))
	subDT[, relative := Normalized_count.y / Normalized_count.x]
	
	p4 <- ggplot(subDT[hsplit.y=='roll',], aes(x=Date, y=relative, colour=vsplit.y)) +
		facet_grid(hsplit.y ~ vsplit.y, scales = 'free_y') +
		geom_hline(yintercept = 1, size=0.15) +
		geom_line(size=0.75) +
		scale_y_continuous(trans='log2', breaks=c(1/32,1/16,1/8,1/4,1/2,1,2,4,8,16,32)) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(1/33, 33)) +
		# annotation_logticks(sides='lr') +
		labs(title=paste(i, 'relative to Austria'), y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					axis.text.y.left = element_text())
	
	print( p1 / (p2a / p2b) / p4)
=======
## Detailed

### Countries of personal interest

#### More plots

```{r, fig.height=20, fig.width=10, warning=FALSE}
for (i in topInterest) {
	more_plots(tidyDT, expDT, rateDT, i, homecountry)
  
  ggsave(paste0('~/covid/', i, '_more.png'), plot = last_plot(), scale = 1,
       width = 20, height = 40, units = "cm", dpi = 300, limitsize = TRUE)
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
}
```

#### Fewer plots

<<<<<<< HEAD
```{r, fig.height=16, fig.width=9}
for (i in c('Switzerland', 'Slovakia', 'Slovenia', 'Czechia', 'Hungary') ) {
	cat(i)
	
	subDT <- melt(DT[, .(Date, Country, norm_new_Cases, norm_new_Deaths, norm_tot_Cases, norm_tot_Deaths, norm_roll_Cases, norm_roll_Deaths)],
								id.vars = c('Date', 'Country'), variable.name = 'Type', value.name = 'Normalized_count')
	subDT[grepl('Death', Type), vsplit := 'Deaths']
	subDT[!grepl('Death', Type), vsplit := 'Cases']
	subDT[, hsplit := sub('_Cases|_Deaths', '', sub('norm_', '', Type), perl=TRUE)]
		
	p1 <- ggplot(subDT, aes(x=Date, y=Normalized_count, group=Country, colour=vsplit)) +
		facet_grid(hsplit ~ vsplit, scales = 'free_y') +
		geom_line(colour='#000000', alpha=0.1, size=0.25) +
		geom_line(data=subDT[Country==i,], size=0.75) +
		scale_y_log10() +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(1,NA)) +
		annotation_logticks(sides='lr') +
		labs(title=paste(i, ': new, total &', W, '-day rolling sum, per', P, 'residents'), y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					panel.grid = element_blank())
	
	p2a <- relative_plot(DT, i, 'total_Cases', 'roll_Cases', case_col)
	p2b <- relative_plot(DT, i, 'total_Deaths', 'roll_Deaths', death_col)

	
	# p3a <- rate_plot(DT, i, 'roll_Cases_Rate', 'Daily change rates', case_col)
	# p3b <- rate_plot(DT, i, 'roll_Deaths_Rate', '', death_col)

	
	subDT <- merge(refDT, mDT[Country==i,], by=c('Date', 'Type'))
	subDT[, relative := Normalized_count.y / Normalized_count.x]
	
	p4 <- ggplot(subDT[hsplit.y=='roll',], aes(x=Date, y=relative, colour=vsplit.y)) +
		facet_grid(hsplit.y ~ vsplit.y, scales = 'free_y') +
		geom_hline(yintercept = 1, size=0.15) +
		geom_line(size=0.75) +
		scale_y_continuous(trans='log2', breaks=c(1/32,1/16,1/8,1/4,1/2,1,2,4,8,16,32)) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(1/33, 33)) +
		# annotation_logticks(sides='lr') +
		labs(title=paste(i, 'relative to Austria'), y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					axis.text.y.left = element_text())
	
	print( p1 / (p2a / p2b) / p4)
=======
```{r, fig.height=15, fig.width=10, warning=FALSE}
for (i in topInterest) {
	fewer_plots(tidyDT[hsplit=='roll',], expDT, rateDT, i, homecountry)
  
  ggsave(paste0('~/covid/', i, '_fewer.png'), plot = last_plot(), scale = 1,
       width = 20, height = 40, units = "cm", dpi = 300, limitsize = TRUE)
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
}
```

## Comparative

### Selected countries

<<<<<<< HEAD
	
	# p3a <- rate_plot(DT, i, 'roll_Cases_Rate', 'Daily change rates', case_col)
	# p3b <- rate_plot(DT, i, 'roll_Deaths_Rate', '', death_col)

	
	subDT <- merge(refDT, mDT[Country==i,], by=c('Date', 'Type'))
	subDT[, relative := Normalized_count.y / Normalized_count.x]
	
	p4 <- ggplot(subDT[hsplit.y=='roll',], aes(x=Date, y=relative, colour=vsplit.y)) +
		facet_grid(hsplit.y ~ vsplit.y, scales = 'free_y') +
		geom_hline(yintercept = 1, size=0.15) +
		geom_line(size=0.75) +
		scale_y_continuous(trans='log2', breaks=c(1/32,1/16,1/8,1/4,1/2,1,2,4,8,16,32)) +
		scale_colour_manual(values=c(Deaths=death_col, Cases=case_col)) +
		coord_cartesian(ylim=c(1/33, 33)) +
		# annotation_logticks(sides='lr') +
		labs(title=paste(i, 'relative to Austria'), y='', x='') +
		theme_bw() + 
		theme(legend.position = 'none',
					axis.text.y.left = element_text())
	
	print( p1 / (p2a / p2b) / p4)
=======
```{r, fig.height=20, fig.width=10, warning=FALSE}
countries <- c(topInterest, other_countries, neighbour_countries)

for (j in c('norm_roll_Cases', 'norm_roll_Deaths', 'norm_new_Tests')) {
	compare_numbers(tidyDT, j, countries)
  
  ggsave(paste0('~/covid/', j, '_comp.png'), plot = last_plot(), scale = 1,
       width = 30, height = 3 * length(countries), units = "cm", dpi = 300, limitsize = TRUE)
>>>>>>> 7b891eca658a5cef042e23b26f96d36e863fe2c8
}

for (j in c('Cases', 'Deaths')) {
  compare_exp(expDT, j, 'norm_Total', 'norm_Roll', countries)
    
  ggsave(paste0('~/covid/', j, '_comp.png'), plot = last_plot(), scale = 1,
       width = 30, height = 3 * length(countries), units = "cm", dpi = 300, limitsize = TRUE)
  }
```


# Session

```{r}
sessionInfo()
```