---
title: "PCA & Correlations"
author: "Kimon Froussios"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
params:
  cts: "./null_xref.tsv"               # a file MUST be specified
  covars: "./deseq2_coldata.txt"       # a file MUST be specified
  pdf: "./null.pdf"                    # a file MUST be specified
  prefix: "./null"                     # for the correlation matrices
  nidcols: 1
  idcol: 1
  minL: 0.75            # Minimum component loading, for selection of features to report in PCA
  ntop: 50              # maximum number of hits to list
editor_options:
  chunk_output_type: console
---

\VignetteEngine{knitr::knitr}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

Parameters

```{r message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(plotly)
library(htmltools)
library(patchwork)
library(matrixStats)
library(ggExtra)
library(ggrepel)

options(scipen=2)

theme_set(theme_bw())

# params <- list(cts=opt$countsFile, covars=opt$samplesFile, nidcols=opt$nidcols, idcol=opt$idcol, pdf=file.path(opt$resultsDir, sub('.txt|.tsv', '_pca.pdf', basename(opt$countsFile))), minL=0.75, ntop=50, prefix=file.path(opt$resultsDir, sub('.txt|.tsv', '', basename(opt$countsFile))) )
print(params)
```

```{r functions}
# Correlations
my_pairwise_internal_corels <- function(mat, samples, method = "pearson", prefix=params$prefix, txs=4) {
  # Correlations
  cormat <- cor(mat, method=method)
  
  # Cluster
  hcfit <- hclust(dist(scale(cormat, center=TRUE)))
  rn <- rownames(cormat)
  cormat <- cormat[samples, samples]                    # Supplied order
  cormat2 <- cormat                                     # Duplicate on which to delete the diagonal with original order.
  cormat3 <- cormat[rn[hcfit$order], rn[hcfit$order]]   # Duplicate in clustered order.
  cormat4 <- cormat3                                    # duplicate on which to delete the diagonal with in clustered order.
  
  # Delete diagonal half for the numeric labels.
  for (r in 1:nrow(cormat2)) {
    for (c in 1:ncol(cormat2)) {
      if (c <= r) {
        cormat2[r, c] <- NA_real_
      }
    }
  }
  for (r in 1:nrow(cormat4)) {
    for (c in 1:ncol(cormat4)) {
      if (c <= r) {
        cormat4[r, c] <- NA_real_
      }
    }
  }
  
  # Restructure for plotting.
  rn <- rownames(cormat)
  cormat <- as.data.table(cormat)
  cormat[, observation1 := factor(rn, ordered=TRUE, levels=rn)]
  cormat <- melt(cormat, id.vars = "observation1", value.name = "Correlation", variable.name = "observation2")
  cormat[, observation2 := factor(observation2, ordered=TRUE, levels=rn)]
  
  rn2 <- rownames(cormat2)
  cormat2 <- as.data.table(cormat2)
  cormat2[, observation1 := factor(rn2, ordered=TRUE, levels=rn2)]
  cormat2 <- melt(cormat2, id.vars = "observation1", value.name = "Correlation", variable.name = "observation2")
  cormat2[, observation2 := factor(observation2, ordered=TRUE, levels=rn2)]
  cormat2 <- cormat2[!is.na(Correlation)]
  
  rn3 <- rownames(cormat3)
  cormat3 <- as.data.table(cormat3)
  cormat3[, observation1 := factor(rn3, ordered=TRUE, levels=rn3)]
  cormat3 <- melt(cormat3, id.vars = "observation1", value.name = "Correlation", variable.name = "observation2")
  cormat3[, observation2 := factor(observation2, ordered=TRUE, levels=rn3)]
  
  rn4 <- rownames(cormat4)
  cormat4 <- as.data.table(cormat4)
  cormat4[, observation1 := factor(rn3, ordered=TRUE, levels=rn4)]
  cormat4 <- melt(cormat4, id.vars = "observation1", value.name = "Correlation", variable.name = "observation2")
  cormat4[, observation2 := factor(observation2, ordered=TRUE, levels=rn4)]
  cormat4 <- cormat4[!is.na(Correlation)]
  
  # Text colour switch for the dynamic range
  m <- min(cormat4$Correlation, na.rm=TRUE)
  M <- max(cormat4$Correlation, na.rm=TRUE)
  colourswitch <- c( m + 0.49 * (M-m),  m + 0.51 * (M-m) )
  
  
  # Square. Custom order. No values. Full range.
  p1 <- ggplot(cormat, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Custom order. No values. Dynamic range.
  p1a <- ggplot(cormat, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "forestgreen" ) +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Triangle. Custom order. With values. Full range.
  p2 <- ggplot(cormat2, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=Correlation >= -0.60 & Correlation <= 0.60 ), size=rel(txs)) +
    scale_x_discrete(position = "top") +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_colour_manual(values=c('FALSE'="black", 'TRUE'="white"), na.value="forestgreen", guide="none") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Triangle. Custom order. With values. Dynamic range.
  p2a <- ggplot(cormat2, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=(Correlation <= colourswitch[2]) ), size=rel(txs)) +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "transparent" ) +
    scale_colour_manual(values=c("black", "white"), na.value="transparent", guide="none") +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Custom order. With values. Full range.
  p12 <- ggplot(cormat, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(data=cormat2, aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=Correlation >= -0.60 & Correlation <= 0.60 ), size=rel(txs)) +
    scale_x_discrete(position = "top") +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_colour_manual(values=c('FALSE'="black", 'TRUE'="white"), na.value="forestgreen", guide="none") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Custom order. With values. Dyhamic range.
  p12a <- ggplot(cormat, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(data=cormat2, aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=(Correlation <= colourswitch[2]) ), size=rel(txs)) +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "transparent" ) +
    scale_colour_manual(values=c("black", "white"), na.value="transparent", guide="none") +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  
  # Square. Clustered order. No values. Full range.
  p3 <- ggplot(cormat3, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Clustered order. No values. Dyhamic range.
  p3a <- ggplot(cormat3, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "forestgreen" ) +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Triangle. Clustered order. with values. Full range.
  p4 <- ggplot(cormat4, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=Correlation >= -0.60 & Correlation <= 0.60 ), size=rel(txs)) +
    scale_x_discrete(position = "top") +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_colour_manual(values=c('FALSE'="black", 'TRUE'="white"), na.value="forestgreen", guide="none") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Triangle. Clustered order. with values. Dyhamic range.
  p4a <- ggplot(cormat4, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=(Correlation <= colourswitch[2]) ), size=rel(txs)) +
    scale_x_discrete(position = "top") +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "transparent" ) +
    scale_colour_manual(values=c("black", "white"), na.value="transparent", guide="none") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Clustered order. With values. Full range.
  p34 <- ggplot(cormat3, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(data=cormat4, aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=Correlation >= -0.60 & Correlation <= 0.60 ), size=rel(txs)) +
    scale_fill_gradientn(limits=c(-1, 1), colors=c("lightskyblue", "dodgerblue3", "darkblue", "black", "darkred", "red", "gold"), na.value = "forestgreen" ) +
    scale_colour_manual(values=c('FALSE'="black", 'TRUE'="white"), na.value="forestgreen", guide="none") +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  # Square. Clustered order. With values. Dyhamic range.
  p34a <- ggplot(cormat3, aes(x=observation1, y=observation2)) +
    geom_tile(aes(fill=Correlation)) +
    geom_text(data=cormat4, aes(label=sub('0.', '.', as.character(round(Correlation, 2))), colour=(Correlation <= colourswitch[2]) ), size=rel(txs)) +
    scale_fill_gradientn(colors=c("black", "red", "gold", "white"), na.value = "forestgreen" ) +
    scale_colour_manual(values=c("black", "white"), na.value="transparent", guide="none") +
    scale_x_discrete(position = "top") +
    labs(x='', y='', title=paste(paste(toupper(substr(method, 1, 1)), tolower(substr(method, 2, nchar(method))), "'s", sep=""), "correlation - Clustered")) +
    theme(axis.text.x=element_text(angle=90, hjust=0, vjust=0.5),
          panel.grid = element_blank() )
  
  fwrite(dcast(cormat2, observation1 ~ observation2, value.var = "Correlation"),
         file=paste0(prefix, '_cor.txt'),
         sep='\t', quote = FALSE, row.names = FALSE, col.names = TRUE)
  
  
  return( list(sfrnc=p1, tfrnc=p2, frnc=p12,
               sdrnc=p1a, tdrnc=p2a, drnc=p12a,
               sfrc=p3, tfrc=p4, frc=p34,
               sdrc=p3a, tdrc=p4a, drc=p34a) )
}

# PCA
# (scaling and centering of genes to one another, NOT samples to one another)
do_pca <- function(countsmat, covars, center = TRUE, scale = TRUE, loadthresh = params$minL, prefix = paste0(params$prefix, '_pc')) {
  # countsmat = counts

  # Gene variances
  genevar <- data.table(name = rownames(countsmat),
                        Mean = rowMeans(countsmat),
                        StDev = rowVars(countsmat) )

  # Rotate counts so genes are variables and samples are observations
  countsmat <- t(countsmat) # rotate 270 so the final rotate for correlations results in the same sample order as the initial correlations

  # Add covariates info
  counts <- cbind(covars, as.data.table(countsmat))

  # Calculate principal components (after getting rid of zero-variance genes that can't be standardized).

  nonconstant <- which(apply(countsmat, 2, var)!=0)
  countsmat <- countsmat[, nonconstant]
  nvars <- dim(countsmat)[2]

  pca <- prcomp(countsmat, center = center, scale. = scale)
  srn <- sqrt(nrow(countsmat) - 1)
  pc <- sweep(pca$x, 2, 1 / (pca$sdev * srn), FUN = '*')
  pc <- cbind(pc, data.frame(sample = rownames(pc)))
  #covars <- cbind(covars, data.frame(sample = rownames(covars)))
  pc <- as.data.frame(merge(pc, covars, by="sample", all = TRUE))
  npc <- sum(pca$sdev > 1)

  # Screeplot.

  pcaimp <- as.data.table(cbind(as.data.frame(t(summary(pca)$importance)),
                  data.frame(PC = 1:length(colnames(pca$x)))))
  pcaimp <- melt(pcaimp, variable.name = "type", value.name = "Proportion", id.vars = c("PC", "Standard deviation"))

  pcmax=15
  pimp <- ggplot(pcaimp[(PC<pcmax),], aes(x=PC, y=Proportion)) +
    facet_grid(type ~ ., scales="free_y") +
    geom_bar(aes(fill=type), stat="identity", width=0.5) +
    geom_text_repel(aes(x=PC-0.4, label=paste0(round(Proportion*100, 1),"%"), colour=rev(type)), ylim=c(0,max(pcaimp[type=='Proportion of Variance', Proportion])), angle=90, vjust=0, size=rel(3)) +
    geom_text_repel(aes(label=paste0(round(Proportion*100, 1),"%"), colour=type), ylim=c(0, max(pcaimp[type=='Cumulative Proportion', Proportion])), angle=90, vjust=0, direction= 'y', size=rel(3)) +
    geom_line(aes(colour=type)) +
    scale_fill_manual(values = c("grey50", "transparent")) +
    scale_colour_manual(values = c("transparent", "grey25")) +
    scale_x_continuous(breaks = seq(1, pcmax, 2)) +
    labs(title = "PCA Screeplot", subtitle=paste0(nvars, " features, ", npc, " PCs"), x = "Principal Component") +
    theme(axis.title.y = element_blank(),
          legend.position = "none")

  # Top loadings for the first 3 PCs.

  loads <- as.data.table(t(t(pca$rotation) * pca$sdev))
  loads[, rowID := rownames(pca$rotation)]
  loads <- melt(loads, value.name = "Loading", id.vars="rowID", variable.name="PC")
  loads[, absload := abs(Loading)]
  lhigh <- loads$absload > loadthresh     # keep only high loadings
  setorder(loads, PC, -absload)
  sel1 <- loads[lhigh > loadthresh & PC=="PC1", rowID]
  sel2 <- loads[lhigh > loadthresh & PC=="PC2", rowID]
  sel3 <- loads[lhigh > loadthresh & PC=="PC3", rowID]
  sel <- unique(loads[lhigh, rowID])

  # Coefficient names.

  ig <- names(covars)
  ig <- ig[! ig %in% c('sample', 'Sample', 'name', 'Name', 'sizeFactor')]

  # Means and Variances of the top loads in the first 3 PCs.

  genevar = data.table(rowID = colnames(countsmat),
                       Mean = colMeans(countsmat),
                       StDev = colSds(countsmat) )
  genevar[, istop:= rowID %in% sel]

  pvar1 <- ggMarginal(ggplot(genevar, aes(x=Mean, y=StDev, label=rowID, colour=istop)) +
      geom_point(shape=16, size=0.8, alpha=0.5) +
      scale_x_log10() +
      scale_colour_manual(values=c("black", "violetred")) +
      scale_y_log10() +
      labs(x="Mean", y="Standard Deviation", title=paste0("All features (", nvars, ")")) +
      theme(legend.position = "none"),
    type = "histogram")

  pvar2 <- ggplot(genevar[(istop),], aes(x=Mean, y=StDev, label=rowID)) +
      geom_point(shape=16, size=0.8, alpha=0.5, colour="purple") +
      labs(x="Mean", y="Standard Deviation",
           title=paste0("High-loading features (", length(sel), ")"), subtitle = paste0("( >", loadthresh, ") in any of the first 3 PCs")) +
      scale_x_log10() +
      scale_y_log10()

  # Correlation of the samples for the selected features only.

  pcor <- my_pairwise_internal_corels(t(countsmat[, sel]), samples=covars$sample, prefix=prefix)

  # Plot first 3 PCs in 2D pairs.
  # Highlight one variable at a time.
  pc12 <- lapply(ig, function(varname) {
    return(
      ggplot(pc, aes_string(x="PC1", y="PC2", label="sample", colour=varname)) +
        geom_point(alpha=0.8, size=rel(1.5)) +
        coord_fixed() +
        labs(x=paste0("PC1 (", round(pcaimp[1,4]*100, 1), "%)"), y=paste0("PC2 (", round(pcaimp[2,4]*100, 1), "%)"))
    )})

  pc13 <- lapply(ig, function(varname) {
    return(
      ggplot(pc, aes_string(x="PC1", y="PC3", label="sample", colour=varname)) +
        geom_point(alpha=0.8, size=rel(1.5)) +
        coord_fixed() +
        labs(x=paste0("PC1 (", round(pcaimp[1,4]*100, 1), "%)"), y=paste0("PC3 (", round(pcaimp[3,4]*100, 1), "%)"))
    )})

  pc32 <- lapply(ig, function(varname) {
    return(
      ggplot(pc, aes_string(x="PC3", y="PC2", label="sample", colour=varname)) +
        geom_point(alpha=0.8, size=rel(1.5)) +
        coord_fixed() +
        labs(x=paste0("PC3 (", round(pcaimp[3,4]*100, 1), "%)"), y=paste0("PC2 (", round(pcaimp[2,4]*100, 1), "%)"))
    )})

  return(list(pca=pca,
              nvars=nvars,
              nPC=npc,
              pimp=pimp,
              load1=sel1, load2=sel2, load3=sel3,
              pc_1_2=pc12, pc_1_3=pc13, pc_3_2=pc32,
              pvar1=pvar1, pvar2=pvar2,
              pcor1=pcor[['frnc']], pcor2=pcor[['drc']]
              ))
}
```

# Input

```{r input}
covars <- fread(params$covars)

counts <- fread(params$cts)
counts <- counts[, c(names(counts)[1:params$nidcols], covars$sample), with=FALSE]    # Only samples specified in covars.
n <- counts[, params$idcol, with=FALSE]
counts <- as.matrix(counts[, (params$nidcols+1):length(counts)])
rownames(counts) <- n[[1]]
counts[is.na(counts)] <- 0
```

# Correlations

```{r correlations, fig.height=16, fig.width=16}
corels <- my_pairwise_internal_corels(counts, samples = covars$sample)
# print(corels[["full"]])
print(corels[["frnc"]])
print(corels[["drc"]])
# ggplotly(corels[["act1"]])
```


# PCA

For this report, the values (ie. expression) of the variables (ie. genes) have been standardized. This compensates for large differences in magnitude range among variables and gives every variable equal weight. Otherwise the variance from high-value variables (ie, high expression genes) would dominate in the determination of principal components. This is typically used when combining variables with different value ranges, such as variables with different measurement units.

NOTE: In the case that the variables are gene expressions, there is some debate on whether high expression genes should be allowed to contribute more than low expression genes or not. My personal judgement and choice is that, although the measurement units are the same across genes, the value ranges are different (some gene products are needed in large quantities, others are effective in small quantities) and therefore standardization is appropriate and preferable.


## Plots

```{r pca}
all_pca <- do_pca(counts, covars)

# print( paste("Num of Variables:", all_pca$nvars, ", Num of Principal Components:", all_pca$nPC) )

# Scree
print(all_pca$pimp)
```

### PC 1 + PC 2

```{r pc12, fig.width=7, fig.height=7}
htmltools::tagList(lapply(all_pca[["pc_1_2"]], function(x) { ggplotly(x) }))
# lapply(all_pca[["pc_1_2"]], function(x) { x })
```

### PC 1 + PC 3

```{r pc13, fig.width=7, fig.height=7}
htmltools::tagList(lapply(all_pca[["pc_1_3"]], function(x) { ggplotly(x) }))
# lapply(all_pca[["pc_1_3"]], function(x) { x })
```

### PC 3 + PC 2

```{r pc32, fig.width=7, fig.height=7}
htmltools::tagList(lapply(all_pca[["pc_3_2"]], function(x) { ggplotly(x) }))
# lapply(all_pca[["pc_3_2"]], function(x) { x })
```

## High loading features

Features with loading >= `r params$minL`.

```{r}
print( all_pca["pvar1"] )
print( all_pca["pvar2"] )
```

## Top high-loading features in the first 3 PCs

Only the `r params$ntop` highest-loading features with loading >= `r params$minL` are listed here, in order of decreasing loading.

### PC1

```{r pc1}
head(all_pca[["load1"]], n=params$ntop)
```

### PC2

```{r pc2}
head(all_pca[["load2"]], n=params$ntop)
```

### PC3

```{r pc3}
head(all_pca[["load3"]], n=params$ntop)
```

## Filtered correlation

Calculated using only the features with loading >= `r params$minL` in any of the first 3 PCs.

```{r fig.width=16, fig.height=16}
print( all_pca[["pcor1"]] )
print( all_pca[["pcor2"]] )
```


# Write plots to PDF

```{r pdf}
## Plots to PDF ##


pdf(params$pdf,
    width=max(12, ceiling(nrow(covars) / 3)), height=max(12, ceiling(nrow(covars) / 3))) # Scale up when many samples, but don't scale down when few.

# Correlation plots
print(corels[['sfrnc']])
print(corels[['tfrnc']])
print(corels[['sdrc']])
print(corels[['tdrc']])

# PCA plots
print(all_pca$pimp)
for (x in all_pca[["pc_1_2"]]) {
  print(x)
}
for (x in all_pca[["pc_1_3"]]) {
  print(x)
}
for (x in all_pca[["pc_3_2"]]) {
  print(x)
}
print(all_pca[["pcor1"]])
print(all_pca[["pcor2"]])

print( ggplot() + geom_point() + theme_void()) # The ggmargin plot in pvar1 doesn't create its own page
print( all_pca["pvar1"] )
print( all_pca["pvar2"] )

dev.off()
```

# Session Info

```{r}
sessionInfo()
```
